### 2015-05-04 ase-2015-tex/abstract.tex
Added separate section for preliminaries, a new figure and explanation for the basic concepts of the symbolic heap, and a new subsection for limitations of the symbolic heap.

A fundamental challenge of using symbolic execution for software
analysis is the treatment of dynamically allocated data. This paper
introduces a compact symbolic heap representation based on a field
sensitive points-to graph with edges annotated with disjunctive
formulas describing aliasing relationships, along with a set of
rewrite rules that describe the initialization, dereferencing, and
update to fields in the symbolic heaps. The symbolic heap supports
arbitrary recursive data structures and captures all possible heaps
that follow a common control flow path up to the point of
execution. An evaluation of a proof-of-concept implementation in the
Java Pathfinder framework is presented to demonstrate the
computational feasibility of the approach over several classical
symbolic execution benchmarks.

>>>>

A fundamental challenge of using symbolic execution for software
analysis is the treatment of dynamically allocated data. \bah{Existing 
techniques either underapproximate the space of possible inputs, 
or are computationally infeasible.} This paper
introduces a compact symbolic heap representation based on a field
sensitive points-to graph with edges annotated with disjunctive
formulas describing aliasing relationships, along with a set of
rewrite rules that describe the initialization, dereferencing, and
update to fields in the symbolic heaps. The symbolic heap supports
arbitrary recursive data structures and captures all possible heaps
that follow a common control flow path up to the point of
execution. An evaluation of a proof-of-concept implementation in the
Java Pathfinder framework is presented to demonstrate the
computational feasibility of the approach over several classical
symbolic execution benchmarks.

### 2015-05-04 ase-2015-tex/conclusion.tex
Added separate section for preliminaries, a new figure and explanation for the basic concepts of the symbolic heap, and a new subsection for limitations of the symbolic heap.

\section{Conclusion}

Although our approach is inspired by the work in the static analysis
domain, the heap representations used by those techniques do not lend
themselves to symbolic execution since they rely on approximations,
they are inflexible in what types of relations can be expressed, and
they rely on rewriting of constraints.  In this work, we present a
heap representation for symbolic execution that maintains exactness,
represents arbitrary heap structures without restrictions on aliasing
relationships, and supports dynamic on-the-fly initialization of heap
structures. This unique combination of capabilities enables the
symbolic heap algorithm to outperform state-of-the-art symbolic
execution techniques, and in some cases, the performance gains are
considerable. We believe that performance gains in symbolic execution
can naturally benefit a variety of analyses based on symbolic
execution, for example, verification of properties, especially those
related to the heap, test case generation, program evolution
techniques such as directed incremental symbolic
execution~\cite{person:pldi2011}, and other back-end analyses. In
future work we plan to evaluate the impact of analyses that rely on
symbolic execution.

>>>>

\section{Discussion}

\subsection{Limitations}
\bah{This paper presents a symbolic heap that is both sound and complete with respect to the properties verifiable through symbolic execution. However, when the symbolic heap is used in conjunction with any analysis method, it is subject the limitations inherent to the underlying technique. For example, for programs that do not exhibit self-limiting behavior, as a practical matter symbolic execution generally requires some sort of bounds to limit infinite loops and recursion. This means that some sort of constraint will invariably be imposed by the symbolic execution engine, impacting the heap as it would any other symbolic value.  In this case, the advantage of our method is that these constraints may be freely chosen to suit the purposes of the analysis at hand, rather than being imposed by a limitation of the heap representation.}

\subsection{Conclusion}

In this work, we present a
heap representation for symbolic execution that maintains exactness,
represents arbitrary heap structures without restrictions on aliasing
relationships, and supports dynamic on-the-fly initialization of heap
structures. This unique combination of capabilities enables the
symbolic heap algorithm to outperform state-of-the-art symbolic
execution techniques, and in some cases, the performance gains are
considerable. We believe that performance gains in symbolic execution
can naturally benefit a variety of analyses based on symbolic
execution, for example, verification of properties, especially those
related to the heap, test case generation, program evolution
techniques such as directed incremental symbolic
execution~\cite{person:pldi2011}, and other back-end analyses. In
future work we plan to evaluate the impact of analyses that rely on
symbolic execution.

### 2015-05-04 ase-2015-tex/evaluation.tex
Added separate section for preliminaries, a new figure and explanation for the basic concepts of the symbolic heap, and a new subsection for limitations of the symbolic heap.

The prototype of our~\symtxt{} algorithm is implemented as a new
extension to SPF.  In addition to the operations presented in the in
this paper, the implementation contains support for operations over
integers.  We are currently working on adding support for floating
point operations, arrays, and bit-operations.  We use the
\texttt{jConstraints} library which provides support for handling heap
constraints~\cite{ase2014-ghilrr,jpf2014-dghirr}. It uses the z3
solver~\cite{deMouraBjorner08Z3} as the underlying constraint
solver. We leverage the incremental solving feature in z3 in our
experiments. We also implement other optimizations that cache and
re-use solutions to previous constraints. These optimizations are an
important part of our implementation because only small portions of
the heap constraint changes from one state to another. 
\bah{The solver optimizations used in the \symtxt{} are not applicable to \gsetxt{}-style techniques, as they do not generate or solve any heap-related constraints. Even with the solver optimizations these techniques would have the same performance characteristics. The overhead in \gsetxt{} arises from case-splitting (generating states) rather than solver time. }

\bah{The benchmarks all include invariants and traversals of data structures. The chosen benchmarks have a high proportion of heap-related operations, putting maximum focus on the heap analysis methods under investigation. They are included to provide evidence of the general computational efficiency of the SH approach. In future work we plan to evaluate it on a more diverse set of benchmarks. The BinarySearchTree test uses the class of the same name from Alan Weiss' book titled ``Data Structures and Algorithm Analysis in Java''. The test method first asserts a repOk() invariant, before conducting a traversal to find an arbitrary integer key in the tree. The TreeMap method comes from the standard java.util library, and is an implementation of a balanced red/black tree. This method asserts a repOk() invariant, before traversing the tree. The LinkedList test starts by asserting a repOk() invariant, which checks for loops in the linked list using Floyd’s cycle-finding algorithm. The linked list is then traversed using a contains() method that attempts to locate an arbitrary reference key in the list.}

\bah{All benchmarks were performed with respect to a given k-bounds. k-bounding limits the length of reference chains to some user-defined integer k. k-bounding was chosen because it provides a fair means of comparison of all the techniques under investigation. More details on k-bounding can be found in~\cite{Deng:2007}.}

>>>>

The prototype of our~\symtxt{} algorithm is implemented as a new
extension to SPF.  In addition to the operations presented in the in
this paper, the implementation contains support for operations over
integers.  We are currently working on adding support for floating
point operations, arrays, and bit-operations.  We use the
\texttt{jConstraints} library which provides support for handling heap
constraints~\cite{ase2014-ghilrr,jpf2014-dghirr}. It uses the z3
solver~\cite{deMouraBjorner08Z3} as the underlying constraint
solver. We leverage the incremental solving feature in z3 in our
experiments. We also implement other optimizations that cache and
re-use solutions to previous constraints. These optimizations are an
important part of our implementation because only small portions of
the heap constraint changes from one state to another. 
The solver optimizations used in the \symtxt{} are not applicable to \gsetxt{}-style techniques, as they do not generate or solve any heap-related constraints. Even with the solver optimizations these techniques would have the same performance characteristics. The overhead in \gsetxt{} arises from case-splitting (generating states) rather than solver time.

The benchmarks all include invariants and traversals of data structures. The chosen benchmarks have a high proportion of heap-related operations, putting maximum focus on the heap analysis methods under investigation. They are included to provide evidence of the general computational efficiency of the SH approach. In future work we plan to evaluate it on a more diverse set of benchmarks. The BinarySearchTree test uses the class of the same name from Alan Weiss' book titled ``Data Structures and Algorithm Analysis in Java''. The test method first asserts a repOk() invariant, before conducting a traversal to find an arbitrary integer key in the tree. The TreeMap method comes from the standard java.util library, and is an implementation of a balanced red/black tree. This method asserts a repOk() invariant, before traversing the tree. The LinkedList test starts by asserting a repOk() invariant, which checks for loops in the linked list using Floyd’s cycle-finding algorithm. The linked list is then traversed using a contains() method that attempts to locate an arbitrary reference key in the list.

All benchmarks were performed with respect to a given k-bounds. k-bounding limits the length of reference chains to some user-defined integer k. k-bounding was chosen because it provides a fair means of comparison of all the techniques under investigation. More details on k-bounding can be found in~\cite{Deng:2007}.

### 2015-05-04 ase-2015-tex/intro.tex
Added separate section for preliminaries, a new figure and explanation for the basic concepts of the symbolic heap, and a new subsection for limitations of the symbolic heap.

The contribution of this work is a non-trivial adaptation of concepts
in abstract interpretation to the domain of symbolic execution. The
result of which, unlike~\gsetxt{}, is the ability to efficiently generate
and update \emph{symbolic heaps} to produce a single path constraint
that is sound and complete up to the point of execution. This is
accomplished by representing the symbolic heap with a field sensitive
points-to graph with edges annotated with disjunctive formulas
describing aliasing relationships. Where~\gsetxt{} branches the heap at every
de-reference, the symbolic heap in this work splits only at reference
comparison operations or field operations that may throw exceptions.

@@

\item\textbf{-} The first sound and complete system for reasoning symbolically about the 
set of input heaps along any valid program path.

>>>>

\bah{The contribution of this work is a method for modeling and solving systems 
of constraints over a symbolic input heap.} The
result of which, unlike~\gsetxt{}, is the ability to efficiently generate
and update \emph{symbolic heaps} to produce a single path constraint
that is sound and complete up to the point of execution. This is
accomplished by representing the symbolic heap with a field sensitive
points-to graph with edges annotated with disjunctive formulas
describing aliasing relationships. Where~\gsetxt{} branches the heap at every
de-reference, the symbolic heap in this work splits only at reference
comparison operations or field operations that may throw exceptions.

@@

\item\textbf{-} A system for reasoning symbolically about the 
set of input heaps along any valid program path that is sound
and complete with respect to symbolic execution.

### 2015-05-04 ase-2015-tex/lazy.tex
Added javalite details to main paper.
Added separate section for preliminaries, a new figure and explanation for the basic concepts of the symbolic heap, and a new subsection for limitations of the symbolic heap.

\section{Generalized Symbolic Execution}
\label{app:gse}
The specialization of Javalite to generalized symbolic execution with
lazy initialization using the symbolic heap is in
\figref{fig:lazyInit} for the initialization and \figref{fig:lazy} for
how the initialization is used. Specifically, the field-access rule in
\figref{fig:lazy} uses initialization, $\rightarrow_I^*$, to be sure
the accessed field is instantiated. Initialization in generalized
symbolic execution never happens for more than one object on any use
of $\rightarrow_I^*$: $\Lambda$ is either empty or contains exactly
one location.

>>>>

\subsection{Generalized Symbolic Execution}
\label{app:gse}
The rules to generalized symbolic execution with
lazy initialization using the symbolic heap are in
\figref{fig:lazyInit} for the initialization and \figref{fig:lazy} for
how the initialization is used. Specifically, the field-access rule in
\figref{fig:lazy} uses initialization, $\rightarrow_I^*$, to ensure
the accessed field is instantiated. Initialization in generalized
symbolic execution never happens for more than one object on any use
of $\rightarrow_I^*$: $\Lambda$ is either empty or contains exactly
one location.

### 2015-05-04 ase-2015-tex/summary-heap.tex
Added separate section for preliminaries, a new figure and explanation for the basic concepts of the symbolic heap, and a new subsection for limitations of the symbolic heap.

\section{Our Approach}
\label{sec:sh}

The symbolic heap algorithm is defined using
Javalite~\cite{saints-MS}. Note that the approach is not restricted to
this language or even Java in general; we use Javalite simply as
mechanism to describe our approach. The small-step semantics of
Javalite, defined as a syntactic machine using a CESK architecture,
are modified to perform symbolic execution on a symbolic heap. The
surface (input) syntax and machine (state) syntax are shown
in~\figref{fig:syntax}. Terminals are in bold face while non-terminals
are italicized. Ellipses indicate zero or more repetitions. Tuples
omit the commas for compactness.

A Javalite program, $\cfgnt{P}$, is a registry of classes, $\mu$, with
a tuple indicating a class, $\cfgnt{C}$, and a method, $\cfgnt{m}$,
where execution starts. The only primitive type in Javalite is
Boolean. For simplicity in presentation, classes have only
non-primitive fields, and methods have a single
parameter. Expressions, $\cfgnt{e}$, include statements, and they use
'$\cfgt{:=}$' to indicate assignment and '$\cfgt{=}$' to indicate
comparison.  The dot-operator for field access is replaced by
'$\cfgt{\$}$', and the dot-operator for method invocation is replaced
by '$\cfgt{@}$' since '.' is reserved in PLT Redex. There is no
explicit return statement; rather, the value of the last expression is
used as the return value. A variable is always indicated by \cfgnt{x}
and a value by \cfgnt{v}. A value can be a reference in the heap,
$\cfgnt{r}$, or any of the special values shown
in~\figref{fig:syntax}(a).  Only type correct programs are given as
input to the machine.

The state of the Javalite machine, $\cfgnt{s}$, includes the program
registry, $\mu$, a pair of functions defining the heap
$\lp\cfgnt{L}\ \cfgnt{R}\rp$, a constraint, $\phi$, defined over
references in the heap, the environment, $\eta$, for local variables,
and the continuation $\cfgnt{k}$. The registry never changes so it is
omitted from the state tuple in the rest of the presentation.  The
more complex structures, such as the environment and heap, are defined
as lists that start with empty, \cfgnt{mt}. These lists are treated as
partial functions so $\eta(\cfgnt{x}) = \cfgnt{r}$ is the reference
$\cfgnt{r}$ associated with variable $\cfgnt{x}$. The notation
$\eta^\prime = \eta[\cfgnt{x} \mapsto \cfgnt{v}]$ defines a new
partial function $\eta^\prime$ that is the same as $\eta$ except that
the variable $\cfgnt{x}$ now maps to $\cfgnt{v}$. The continuation
$\cfgnt{k}$ indicates where the hole, $\cfgt{*}$, is located, stores
temporary computation, and keeps track of the next computation. For
example, the continuation $\lp\cfgt{*}\ \cfgt{\$}\ \cfgnt{f}
\rightarrow \cfgnt{k}\rp$ indicates that the machine is evaluating the
expression for the object reference on which the field $\cfgnt{f}$ is
going to be accessed. Once the field access is complete, the machine
continues with $\cfgnt{k}$.

@@

Our symbolic heap is a field sensitive points-to graph with edges
annotated with disjunctive formulas describing aliasing
relationships. We represent the symbolic heap as a bipartite graph on
references, $\cfgnt{r}$, and constraint location pairs,
$\lp\phi\ \cfgnt{l}\rp$. The graph is expressed in the location map,
$\cfgnt{L}$, and the reference map, $\cfgnt{R}$. As done with the
environment, $\cfgnt{L}$ and $\cfgnt{R}$ are treated as partial
functions where $\cfgnt{L}(r) = \{(\phi\ \cfgnt{l})\ ...\}$ is the set
of constraint-location pairs in the heap associated with the given
reference, and $\cfgnt{R}(\cfgnt{l},\cfgnt{f}) = \cfgnt{r}$ is the
reference associated with the given location-field pair in the
heap. References in the graph are immutable. Locations, however, are
mutated on updates.  Immutability of references means that look-ups in
the heap are non-recursive and updates do not modify the current path
condition.

The soundness and completeness proofs in Section~\ref{sec:bisim} rely
on certain characteristics of the symbolic heap that are preserved by
the small-step semantics during symbolic execution:
%\emph{reference partitions}, \emph{null and uninitialized}, \emph{immutability}, \emph{determinism}, and
%\emph{type consistency}.
\begin{compactdesc}
%<<<<<<< HEAD
\item[Reference partitions:] Each reference belongs to one of three partitions: $\mathrm{init}_\cfgnt{r}()$ for the \emph{input
  heap}% that includes $\cfgnt{r}_\mathit{null}$ (these are created by lazy initialization and are the only references that appear as variables in constraints)
  ; $\mathrm{fresh}_\cfgnt{r}()$ for \emph{auxiliary
  literals}% that includes $\cfgnt{r}_\mathit{un}$ (these add structure but do not appear elsewhere)
  ; and $\mathrm{stack}_\cfgnt{r}()$ for \emph{stack
  literals} %(these appear in environments, expressions, and continuations).
\item[Null and Uninitialized:] Every symbolic heap contains 
two special locations, null ($\cfgnt{l}_\mathit{null}$), and uninitialized,
%=======
%\item[Reference partitions] means that a reference belongs to one of three partitions: $\mathrm{init}_\cfgnt{r}()$ for the \emph{input
%  heap} that includes $\cfgnt{r}_\mathit{null}$ (these are created by lazy initialization and are the only references that appear as variables in
%  constraints); $\mathrm{fresh}_\cfgnt{r}()$ for \emph{auxiliary
%  literals} that includes $\cfgnt{r}_\mathit{un}$ (these add structure but do not appear elsewhere); and $\mathrm{stack}_\cfgnt{r}()$ for \emph{stack
%  literals} (these appear in environments, expressions, and continuations).
%\item[Null and Uninitialized] means that every symbolic heap contains 
%a special location for null ($\cfgnt{l}_\mathit{null}$), and a special
%location for uninitialized
%>>>>>>> 49c83f34a0b5afd58adc4b6649b15e2cbc93437c
  ($\cfgnt{l}_\mathit{un}$) with $\cfgnt{L}(\cfgnt{r}_\mathit{null}) =
  \{(\cfgt{true}\ \cfgnt{l}_\mathit{null})\}$ and
  $\cfgnt{L}(\cfgnt{r}_\mathit{un}) =
  \{(\cfgt{true}\ \cfgnt{l}_\mathit{un})\}$. 
\item[Immutability:] References do not change once added to the
location map, eliminating any need for constraint rewriting as the program state evolves.
\item[Determinism:] A reference in $(\cfgnt{L}\ \cfgnt{R})$ cannot point to multiple locations simultaneously: 
$
\forall \cfgnt{r} \in \cfgnt{L}^\leftarrow\ (\forall (\phi\ \cfgnt{l}),(\phi^\prime\ \cfgnt{l}^\prime) \in \cfgnt{L}(r)\ (
(\cfgnt{l} \neq \cfgnt{l}^\prime \vee \phi \neq \phi^\prime) \Rightarrow (\phi \wedge \phi^\prime = \cfgt{false}))
$
where $\cfgnt{L}^\leftarrow$ is the pre-image of $\cfgnt{L}$. 
\item[Type consistency:] All locations associated with a reference have the same type:
$\forall \cfgnt{r} \in \cfgnt{L}^\leftarrow\ (\forall (\phi\ \cfgnt{l}),(\phi^\prime\ \cfgnt{l}^\prime) \in \cfgnt{L}(r)\ (
(\mathrm{Type}\lp\cfgnt{l}\rp = \mathrm{Type}\lp\cfgnt{l}^\prime\rp)))$

@@

\subsection{Semantics of Javalite}

@@

\figref{fig:javalite-common} defines
the rewrite relations for the portion of the Javalite language
semantics that are common to both generalized symbolic execution and
the symbolic heap algorithm. The relation $\rcom$ includes all of these
rules. Excepting \textrm{N{\footnotesize EW}}, the rules do not update
the heap, and are largely concerned with argument evaluation in an
expected way. It is assumed that only type safe programs are input to
the machine so there is no type checking or error conditions. The
machine simply halts if no rewrite is enabled.
\bah{
\input{lazy}	
}
\input{precise-fig}

>>>>

\section{Prelimiaries}
\label{sec:sh}

@@

Our symbolic heap is a field sensitive points-to graph with edges
annotated with disjunctive formulas describing aliasing
relationships. We represent the symbolic heap as a bipartite graph on
references, $\cfgnt{r}$, and constraint location pairs,
$\lp\phi\ \cfgnt{l}\rp$. The graph is expressed in the location map,
$\cfgnt{L}$, and the reference map, $\cfgnt{R}$. As done with the
environment, $\cfgnt{L}$ and $\cfgnt{R}$ are treated as partial
functions where $\cfgnt{L}(r) = \{(\phi\ \cfgnt{l})\ ...\}$ is the set
of constraint-location pairs in the heap associated with the given
reference, and $\cfgnt{R}(\cfgnt{l},\cfgnt{f}) = \cfgnt{r}$ is the
reference associated with the given location-field pair in the
heap. References in the graph are immutable. Locations, however, are
mutated on updates.

\begin{figure}[t]
\begin{center}
\includegraphics[scale=0.2]{../figs/simple_heap_scratch.png}
\end{center}
\caption{Example Heap (scratch image, not for publication!)}
\label{fig:exampleHeap}
\end{figure}

\bah{\figref{fig:exampleHeap} shows an example symbolic heap. References are 
represented by circles and locations are represented by squares. Arrows leaving
references correspond to the value sets returned by the $\cfgnt{L}$ function, and 
arrows leaving the squares correspond to the $\cfgnt{R}$ function.}

The soundness and completeness proofs in Section~\ref{sec:bisim} rely
on certain characteristics of the symbolic heap that are preserved by
the small-step semantics during symbolic execution:

\begin{compactdesc}
\item[Reference partitions:] Each reference belongs to one of three partitions: $\mathrm{init}_\cfgnt{r}()$ for the \emph{input
  heap}% that includes $\cfgnt{r}_\mathit{null}$ (these are created by lazy initialization and are the only references that appear as variables in constraints)
  ; $\mathrm{fresh}_\cfgnt{r}()$ for \emph{auxiliary
  literals}% that includes $\cfgnt{r}_\mathit{un}$ (these add structure but do not appear elsewhere)
  ; and $\mathrm{stack}_\cfgnt{r}()$ for \emph{stack
  literals} %(these appear in environments, expressions, and continuations).
\item[Null and Uninitialized:] Every symbolic heap contains 
two special locations, null ($\cfgnt{l}_\mathit{null}$), and uninitialized,
  ($\cfgnt{l}_\mathit{un}$) with $\cfgnt{L}(\cfgnt{r}_\mathit{null}) =
  \{(\cfgt{true}\ \cfgnt{l}_\mathit{null})\}$ and
  $\cfgnt{L}(\cfgnt{r}_\mathit{un}) =
  \{(\cfgt{true}\ \cfgnt{l}_\mathit{un})\}$. 
\item[Immutability:] References do not change once added to the
location map, eliminating any need for constraint rewriting as the program state evolves.
\item[Determinism:] A reference in $(\cfgnt{L}\ \cfgnt{R})$ cannot point to multiple locations simultaneously: 
$
\forall \cfgnt{r} \in \cfgnt{L}^\leftarrow\ (\forall (\phi\ \cfgnt{l}),(\phi^\prime\ \cfgnt{l}^\prime) \in \cfgnt{L}(r)\ (
(\cfgnt{l} \neq \cfgnt{l}^\prime \vee \phi \neq \phi^\prime) \Rightarrow (\phi \wedge \phi^\prime = \cfgt{false}))
$
where $\cfgnt{L}^\leftarrow$ is the pre-image of $\cfgnt{L}$. 
\item[Type consistency:] All locations associated with a reference have the same type:
$\forall \cfgnt{r} \in \cfgnt{L}^\leftarrow\ (\forall (\phi\ \cfgnt{l}),(\phi^\prime\ \cfgnt{l}^\prime) \in \cfgnt{L}(r)\ (
(\mathrm{Type}\lp\cfgnt{l}\rp = \mathrm{Type}\lp\cfgnt{l}^\prime\rp)))$

@@

\subsection{Semantics}

In this work, symbolic execution semantics are is defined using
a small-step operational semantics similar to the one presented 
in~\cite{saints-MS}. The small-step semantics,
defined as a syntactic machine using a CESK architecture~\cite{Felleisen:1992},
are modified to perform symbolic execution on a symbolic heap. The
surface (input) syntax and machine (state) syntax are shown
in~\figref{fig:syntax}. Terminals are in bold face while non-terminals
are italicized. Ellipses indicate zero or more repetitions. Tuples
omit the commas for compactness.

A program, $\cfgnt{P}$, is a registry of classes, $\mu$, with
a tuple indicating a class, $\cfgnt{C}$, and a method, $\cfgnt{m}$,
where execution starts. The only primitive type in Javalite is
Boolean. For simplicity in presentation, classes have only
non-primitive fields, and methods have a single
parameter. Expressions, $\cfgnt{e}$, include statements, and they use
'$\cfgt{:=}$' to indicate assignment and '$\cfgt{=}$' to indicate
comparison.  The dot-operator for field access is replaced by
'$\cfgt{\$}$', and the dot-operator for method invocation is replaced
by '$\cfgt{@}$' since '.' is reserved in PLT Redex. There is no
explicit return statement; rather, the value of the last expression is
used as the return value. A variable is always indicated by \cfgnt{x}
and a value by \cfgnt{v}. A value can be a reference in the heap,
$\cfgnt{r}$, or any of the special values shown
in~\figref{fig:syntax}(a).  Only type correct programs are given as
input to the machine.

The state of the machine, $\cfgnt{s}$, includes the program
registry, $\mu$, a pair of functions defining the heap
$\lp\cfgnt{L}\ \cfgnt{R}\rp$, a constraint, $\phi$, defined over
references in the heap, the environment, $\eta$, for local variables,
and the continuation $\cfgnt{k}$. The registry never changes so it is
omitted from the state tuple in the rest of the presentation.  The
more complex structures, such as the environment and heap, are defined
as lists that start with empty, \cfgnt{mt}. These lists are treated as
partial functions so $\eta(\cfgnt{x}) = \cfgnt{r}$ is the reference
$\cfgnt{r}$ associated with variable $\cfgnt{x}$. The notation
$\eta^\prime = \eta[\cfgnt{x} \mapsto \cfgnt{v}]$ defines a new
partial function $\eta^\prime$ that is the same as $\eta$ except that
the variable $\cfgnt{x}$ now maps to $\cfgnt{v}$. The continuation
$\cfgnt{k}$ indicates where the hole, $\cfgt{*}$, is located, stores
temporary computation, and keeps track of the next computation. For
example, the continuation $\lp\cfgt{*}\ \cfgt{\$}\ \cfgnt{f}
\rightarrow \cfgnt{k}\rp$ indicates that the machine is evaluating the
expression for the object reference on which the field $\cfgnt{f}$ is
going to be accessed. Once the field access is complete, the machine
continues with $\cfgnt{k}$.

@@

\figref{fig:javalite-common} defines
the rewrite relations for the portion of the language
semantics that are common to both generalized symbolic execution and
the symbolic heap algorithm. The relation $\rcom$ includes all of these
rules. Excepting \textrm{N{\footnotesize EW}}, the rules do not update
the heap, and are largely concerned with argument evaluation in an
expected way. It is assumed that only type safe programs are input to
the machine so there is no type checking or error conditions. The
machine simply halts if no rewrite is enabled.

\input{lazy}

\input{precise-fig}
